Figures to insert in text

### 2018-03-06 Chapters/Chapter02.tex
SDM

\section{Convergence }

One particular analysis of interest is that of the distance read at a point $\alpha$. Suppose an SDM is trying to read an item written at $\alpha$, but the cues received so far lead to a point of distance $d$ from $\alpha$.  As one reads at $\alpha+d$, a new bitstring $\beta$ is obtained, leading to our question: what is the new distance from $\alpha$ to $\beta$? Is it smaller or larger than $d$? That, of course, depends on the ratio between $d$ and the number of dimensions of the memory.  As we have found out, there are some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.

\subsection{Some initial anomalous results}

As we ran the simulations reflecting some of Kanerva's graphs, one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model, and we propose that this is due to interaction effects between different attractors.

\citet{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Fig. \ref{kanerva-table-7-2}). The original prediction considered that the read distance would decline when inside the critical distance in increase afterwards, converging to a \textasciitilde 500-bit distance.  At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance point.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
\label{kanerva-table-7-2}}
\end{figure}

Our preliminary results show that the theoretical prediction is not accurate.  There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Fig. \ref{sdm-10000w-table-7-2}). Our results were obtained using a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, the same used by Kanerva.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/sdm-10000w-table-7-2.png}
\caption{Results generated by the framework diverging from Kanerva's original Table 7.2. Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
\label{sdm-10000w-table-7-2}}
\end{figure}

But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Fig. \ref{sdm-100w-table-7-2}). This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
\label{sdm-100w-table-7-2}}
\end{figure}

To obtain the results from Fig. \ref{sdm-10000w-table-7-2} and \ref{sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin. Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM. Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000. Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig. \ref{sdm-10000w-notX-table-7-2}). Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.

@@

@@

Let us revisit, in these concluding thoughts, the emphasis employed over speed of computation.  At first sight, that might seem like a typical objective of efficiency in computer science. But we are not only interested in the computer science effects here --- the ambition is different. More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.

If

>>>>

\section{Convergence of $\eta_x$ to $\eta$}

One particular analysis of Kanerva's interest is that of the distance read at a point $\eta_x$. Suppose an SDM is trying to read an item written at $\eta$, but the cues received so far lead to a point of distance $x$ from $\eta$.  As one reads at $\eta_x$, a new bitstring $\beta$ is obtained, leading to Kanerva's question: what is the new distance from $\eta$ to $\beta$? Is it smaller or larger than $x$? That, of course, depends on the ratio between $x$ and the number of dimensions of the memory.

\citet[p.70]{Kanerva1988} originally predicted a \textasciitilde 500-bit distance after a point (Figure \ref{fig:kanerva-figure-7.3}). The original prediction considered that the read distance would decline when inside the critical distance and increase afterwards, converging to a \textasciitilde 500-bit distance.  At this point, each read would lead to a different, orthogonal, \textasciitilde 500-bit distance bitstring. He analyzed specifically an SDM with 1,000 bits and 10,000 random bitstrings written into it.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/kanerva-table-7-2-original.png}
\caption{Kanerva's original Figure 7.3 (p. 70) predicting a \textasciitilde 500-bit distance after a point.
\label{fig:kanerva-figure-7.3}}
\end{figure}

As we ran the simulations, this one in particular struck our attention: The new distances obtained after a read operation were not perfectly predicted by the theoretical model.
%and we propose that this is due to interaction effects between different attractors.

We have strictly followed Kanerva's configuration and, even so, we have found out some deviations from Kanerva's original theoretical analysis and the results obtained by simulation.

In details, we have created a SDM with $n=1,000$, $H=1,000,000$, and $r=451$. Then, we have generated 10,000 random bitstrings and written them into the memory. Then, we have generated a reference bitstring (bs\_ref) and written it into the memory. Then, we have executed the following steps with $x$ from 0 to 1,000: (i) copy bs\_ref into a new bitstring; (ii) randomly flipped $x$ bits of the copy; (iii) read from the memory in the copy address; and (iv) stored the distance between the returned bitstring and bs\_ref. Finally, we have plotted Figure \ref{fig:sdm-10000w-table-7-2}.

\begin{figure}[h]
\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2.png}
\caption{Results generated by the framework diverging from Kanerva's original Table 7.2. Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
\label{fig:sdm-10000w-table-7-2}}
\end{figure}

Figure \ref{fig:sdm-10000w-table-7-2} has a lot of noise because we have read only once for each distance $x$ and Kanerva has predicted the average distance. So, we have changed the steps to run $k$ reads and store the average new distance. We run with $k=6$, and the results can be seen in Figure \ref{fig:sdm-10000w-table-7-2-6-samples}, which has a way lower noise and still holds the divergence.

\begin{figure}[h]
\centering\includegraphics[width=\textwidth]{images02/sdm-10000w-table-7-2-6-samples.png}
\caption{Results generated by the framework diverging from Kanerva's original Table 7.2. Here we had a 1,000 bit, 1,000,000 hard-location SDM with exactly 10,000 random bitstrings written into it, which was also Kanerva's configuration.
\label{fig:sdm-10000w-table-7-2-6-samples}}
\end{figure}

Our results show that the theoretical prediction is not accurate.  There are interaction effects from one or more of the attractors created by the 10,000 writes, and these attractors seem to raise the distance beyond \textasciitilde 500 bits (Figure \ref{fig:sdm-10000w-table-7-2}).

But, when we reduced the number of random bitstrings written in the SDM from 10,000 to only 100, the results reflected very well the Kanerva's theoretical expectation (Fig. \ref{fig:sdm-100w-table-7-2}). This result strengthens our hypothesis that the disparities in the computational results are due to the interaction effect of high numbers of different attractors.

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
\label{fig:sdm-100w-table-7-2}}
\end{figure}

To obtain the results from Figures \ref{fig:sdm-10000w-table-7-2} and \ref{fig:sdm-100w-table-7-2}, we had to write 10,000 random bitstrings to an SDM, and then randomly choose one of those bitstrings to be our origin. Finally, we randomly flipped some bits from the origin bitstring and executed a reading operation in the SDM. Thereby, in order to show the interaction effects more clearly, we wrote a handmade bitstring to the SDM which had all bits inverted in relation to the origin bitstring --- their hamming distance was equal to 1,000. Our handmade bitstring was acting as an opposite attractor, and one can see the accelerating effects towards convergence to both attractors: the origin and the handmade bitstrings (Fig. \ref{sdm-10000w-notX-table-7-2}). Here we had the exact same configuration of Figure \ref{sdm-10000w-table-7-2}, with the addition of the single opposite attractor.

@@

\chapter{Results (ii): Loss of neurons}

\chapter{Results (iii): Generalized read operation}

@@

Let us revisit, in this concluding thoughts, the emphasis employed over speed of computation.  At first sight, that might seem like a typical objective of efficiency in computer science. But we are not only interested in the computer sciece effects here --- the ambition is different. More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.
